We present a convex formulation of compliant contact along with with a robust
iterative method with guaranteed convergence to solve it in practice. Our
formulation eliminates contact constraints analytically to write an
unconstrained convex formulation of frictional contact on velocities only. We
show our method warms-starts effectively, enabling the simulation of complex
multibody systems with frictional contact at interactive rates. 

We provide a succinct physical interpretation of the convex approximation of
contact not previously available in the literature. This allows us to
incorporate not only point contact models of compliance, but also more
sophisticated models of compliant surface patches. Accurate physical modeling of
contact is critical for meaningful sim-to-real transfer. Moreover, we provide
explicit analytical expressions for the contact forces as a function of state,
which allow us to characterize artifacts introduced by the convex approximation
in a novel and transparent way. 

Our time stepping approach is based on the $\theta\text{-method}$, which
includes not only the popular symplectic Euler method but also the second order
midpoint rule. We demonstrate that the midpoint rule can achieve second order
accuracy even in simulation through frictional contact. Moreover, unlike the
symplectic Euler scheme, the midpoint rule is unconditionally stable, enabling
the simulation of mechanical systems with stiff force elements.

We introduce a number of accuracy metrics and show that our method outperforms
existing commercial and open source alternatives without sacrificing accuracy.

Finally, we demonstrate the effectiveness of our approach with the simulation of
a complex dual arm manipulation task with dexterous hands. Our method can
resolve the tight coupling between motions and contact forces with high mass
ratios, which is essential for the stable simulation of manipulation tasks.

Our method is implemented in the open source Drake robotics toolkit.

