\section{Discrete Time Formulation}
\label{sec:discrete_time_formulation}

Our discrete-time model is based on the $\theta\text{-method}$ \cite[\S
II.7]{bib:hairer2008solving} and incorporates the symplectic midpoint rule to
attain second order accuracy and energy conservation. While most of the work in
the literature uses first order time-stepping schemes, the extension to the
second-order midpoint rule is analyzed in \cite{bib:potra2006linearly}. While
the work in \cite{bib:potra2006linearly} uses a polyhedral approximation of the
friction cone that leads to an LCP formulation, our approach does not
approximate the friction cone but introduces the convex approximation of contact
from \cite{bib:anitescu2006} \RedHighlight{TODO: Introduce definition of LCP in
the introduction}. We remark that combining the $\theta\text{-method}$ with the
convex approximation of contact is novel to our work. 

Time is discretized into intervals of fixed size $dt$ and we seek to advance the
state of the system from time $t^n$ to the next step at $t^{n+1} = t^n + dt$. In
the $\theta\text{-method}$ variables are evaluated at intermediate time steps
$t^\theta = \theta t^{n+1}+(1-\theta)t^{n}$, with $\theta \in [0, 1]$. We define
\emph{mid-step quantities} $\mf{q}^{\theta_{q}}$, $\mf{v}^{\theta_{v}}$, and
$\mf{v}^{\theta_{vq}}$ in accordance with the standard $\theta\text{-method}$
using scalar parameters $\theta_q$ $\theta_v$ and $\theta_{vq}$
\begin{eqnarray}
	\mf{q}^{\theta_q} &=& \theta_q\mf{q} + (1-\theta_{q})\mf{q}_0,\nonumber\\
	\mf{v}^{\theta_v} &=& \theta_v\mf{v} + (1-\theta_v)\mf{v}_0,\nonumber\\
	\mf{v}^{\theta_{vq}} &=& \theta_{vq}\mf{v} + (1-\theta_{vq})\mf{v}_0.
	\label{eq:theta_method}
\end{eqnarray}

Using the above definitions, we write our discrete update in the following form
\begin{eqnarray}
	\mf{M}(\mf{q}^{\theta_{q}})(\mf{v}-\mf{v}_0)  &=& \nonumber\\
	dt\,\mf{F}(\mf{q}^{\theta_{q}}, \mf{v}^{\theta_v}) &+&
	dt\,\mf{G}(\mf{q}^{\theta_{q}}, \mf{v}^{\theta_v}) +
	\mf{J}(\mf{q}^{\theta_{q}})^T\mf{\bgamma}, \label{eq:v_update}\\
	\mf{q} = \mf{q}_0 + dt \dot{\mf{q}}^{\theta_{vq}} &=& \mf{q}_0 + dt\mf{N}(\mf{q}^{\theta_{q}})\mf{v}^{\theta_{vq}},
	\label{eq:q_update}
\end{eqnarray}
where to simplify notation $(\mf{q}_0, \mf{v}_0)$ denotes the state at $t^n$ and
$(\mf{q}, \mf{v})$ the state at $t^{n+1}$. $\mf{N}(\mf{q})$ is the kinematic map
defined in Eq. (\ref{eq:kinematic_map}). Notice that in the update for $\mf{q}$
we allow for the evaluation of $\mf{N}$ at $t^{\theta_{q}}$ and velocities
at $t^{\theta_{vq}}$. This will allow us to write fully implicit as well as
semi-implicit symplectic schemes within the same framework. For the common case
when $\mf{N}$ is the identity matrix, $\dot{\mf{q}}$ is evaluated at
$t^{\theta_{vq}}$, justifying the notation. Note that using Eq.
(\ref{eq:q_update}) the mid-step configuration $\mf{q}^{\theta_q}$ can be
written as
\begin{equation}
	\mf{q}^{\theta_q}(\mf{v}) = \mf{q}_0 + dt\theta_q\mf{N}(\mf{q}^{\theta_{q}})\mf{v}^{\theta_{vq}}(\mf{v})
	\label{eq:position_update}
\end{equation}


We split forces into two contributions so that $\partial \mf{F}/\partial\mf{q}$ and $\partial \mf{F}/\partial\mf{v}$ are positive definite matrices while the same is generally not true for the gradients of $\mf{G}$. Therefore in $\mf{F}(\mf{q}, \mf{v})$ we will include modeling elements such as spring and dampers and even internal forces
for the modeling of soft-body deformation. $\mf{G}(\mf{q}, \mf{v})$ will include
all other contributions that cannot guarantee positive definiteness of the
gradients such as Coriolis and gyroscopic forces arising in multibody dynamics with generalized coordinates.

The discrete update in Eqs. (\ref{eq:v_update})-(\ref{eq:q_update}) generalizes
some of the most popular schemes for forward dynamics:
\begin{itemize}
	\item Explicit Euler with $\theta_q=\theta_{v}=\theta_{vq} = 0$,
	\item Symplectic Euler with $\theta_{q} = 0$ and $\theta_v = \theta_{vq}=1$,
	\item Implicit Euler with $\theta_{q} = \theta_v = \theta_{vq}= 1$,
	\item Symplectic midpoint rule, which is second order, with $\theta_{q} =
	\theta_v = \theta_{vq}= 1/2$,
\end{itemize}

Similar to the work in \cite{bib:duriez2005realistic} for the simulation of
deformable objects and to projection methods used in fluid mechanics
\cite{bib::bell1991efficient}, we solve Eqs. (\ref{eq:v_update}) and
(\ref{eq:q_update}) in two stages; first we solve for the \emph{free motion
velocities} the system would have in the absence of contact in
\ref{sec:free_motion_velocities} and second, we update these free motion
velocities so that they satisfy conservation of momentum along with contact
constraints in \ref{sec:contact_constrained_velocities}.

\subsection{Free Motion Velocities}
\label{sec:free_motion_velocities}
We first define the momentum quantity
\begin{multline}
	\mf{m}(\mf{v}) =
	\mf{M}(\mf{q}^{\theta_{q}}(\mf{v}))(\mf{v}-\mf{v}_0) -\\
	dt\,\mf{F}(\mf{q}^{\theta_{q}}(\mf{v}), \mf{v}^{\theta_v}(\mf{v}))-
	dt\,\mf{G}(\mf{q}^{\theta_{q}}(\mf{v}), \mf{v}^{\theta_v}(\mf{v}))
	\label{eq:m_definition}
\end{multline}

In the first stage we solve for the
\textit{free motion} velocities $\mf{v}^*$ in the absence of constraint impulses
\begin{eqnarray}
	\mf{m}(\mf{v}^*) &=& \mf{0},
	\label{eq:vstar_definition}\\
	\mf{q}^{\theta_q}(\mf{v}^*) &=& \mf{q}_0 + dt\theta_q\dot{\mf{q}}^{\theta_{vq}}(\mf{v}^*),
	\label{eq:qstar_definition}
\end{eqnarray}

For the implicit Euler scheme and the midpoint rule, Eqs.
(\ref{eq:vstar_definition})-(\ref{eq:qstar_definition}) are implicit in
$\mf{v}^*$ and we solve them using Newton's method. The explicit and symplectic
Euler schemes require inversion of the mass matrix $\mf{M}$, which can be
accomplished efficiently using the $\mathcal{O}(n)$
\emph{Articulated Body Algorithm}
\cite{bib:featherstone2008_rigid_body_dynamics_algorithms}.

For the second stage we linearize $\mf{m}(\mf{v})$ around $\mf{v} = \mf{v}^*$ to write an approximation of Eq. (\ref{eq:v_update}) as
\begin{eqnarray}
	\mf{A}(\mf{v}-\mf{v}^*)= \mf{J}^T\mf{\gamma}
	\label{eq:momentum_linearized}
\end{eqnarray}
with
\begin{eqnarray}
	\mf{A}&=&\mf{M}+dt^2\,\theta_q\theta_{qv}\mf{K}+dt\,\theta_v\mf{D},
	\label{eq:expression_for_A}\\
	\mf{K}(\mf{q}, \mf{v})&=&-\frac{\partial \mf{F}(\mf{q}, \mf{v})}{\partial
	\mf{q}}\frac{\partial\dot{\mf{q}}^{\theta_{vq}}}{\partial\mf{v}},\\
		\mf{D}(\mf{q}, \mf{v})&=&-\frac{\partial \mf{F}(\mf{q},
		\mf{v})}{\partial \mf{v}},
		\label{eq:dissipation_matrix}
\end{eqnarray}
where $\mf{A} \succ 0$ is an approximation to the gradient of $\mf{m}(\mf{v})$. Equation (\ref{eq:momentum_linearized}) is a first order approximation of Eq. (\ref{eq:v_update}) as stated in the following
\begin{theorem}	
	In the absence of impacts Eq. (\ref{eq:momentum_linearized}) is a first
	order approximation of Eq. (\ref{eq:v_update}). That is
	\begin{eqnarray}
		\mf{m}(\mf{v}) = \mf{A}(\mf{v}-\mf{v}^*) + \mathcal{O}(dt^2) =
		\mf{J}^T\mf{\bgamma}
	\end{eqnarray}
\end{theorem}

\begin{IEEEproof}
The Taylor expansion of $\mf{m}(\mf{v})$ at $\mf{v}=\mf{v}^*$ reads
\begin{eqnarray}
	\mf{m}(\mf{v}) &=& \mf{m}^* +
	\nabla\mf{m}^*(\mf{v}-\mf{v}^*) + \mathcal{O}_m(\Vert\mf{v}-\mf{v}^*\Vert^2)\nonumber\\
	&=&\nabla\mf{m}^*(\mf{v}-\mf{v}^*) + \mathcal{O}_m(\Vert\mf{v}-\mf{v}^*\Vert^2)
	\label{eq:m_taylor_expansion}
\end{eqnarray}
where the superscript $^*$ denotes quantities evaluated at $\mf{v}^*$ and we use
the fact that by definition $\mf{m}^*=\mf{m}(\mf{v}^*)=\mf{0}$. The operator
$\nabla$ denotes the gradient with respect to $\mf{v}$.

We evaluate first the gradient of the mass matrix term in Eq.
(\ref{eq:m_definition}) using the chain rule through Eq.
(\ref{eq:position_update})
\begin{eqnarray}
	\nabla[\mf{M}(\mf{q}^{\theta_{q}}(\mf{v}))(\mf{v}-\mf{v}_0)]\Bigr|_{\mf{v}^*}
	= \mf{M}(\mf{q}^{\theta_{q}}(\mf{v}^*)) + \mf{E}^*,
\end{eqnarray}
where we defined
\begin{eqnarray}
	\mf{E}^* =
	dt\theta_q\frac{\partial[\mf{M}(\mf{q}^{\theta_{q}})(\mf{v}^*-\mf{v}_0)]}{\partial\mf{q}}\frac{\partial\dot{\mf{q}}^{\theta_{vq}}}{\partial\mf{v}}.
	\label{eq:E_definition}
\end{eqnarray}

Equation (\ref{eq:E_definition}) involves cumbersome derivatives of the mass matrix with respect to configurations. However, we can still see that $\mf{E}^*=\mathcal{O}(dt\Vert\mf{v}^*-\mf{v}_0\Vert)$ where we are using the big O notation to describe its limiting behavior as $dt\rightarrow 0$. Since the free motion dynamics that defines $\mf{v}^*$ is continuous, we know that $\Vert\mf{v}^*-\mf{v}_0\Vert = \mathcal{O}(dt)$ and therefore $\mf{E}^* = \mathcal{O}(dt^2)$.

We proceed similarly to expand the gradient of $\mf{F}(\mf{v})=\mf{F}(\mf{q}^{\theta_{q}}(\mf{v}), \mf{v}^{\theta_v}(\mf{v}))$ as
\begin{eqnarray}
	\nabla\mf{F} =
	-dt\,\theta_q\theta_{vq}\mf{K}(\mf{q}^{\theta_q}, \mf{v}^{\theta_v})-\theta_v\mf{D}(\mf{q}^{\theta_q}, \mf{v}^{\theta_v})
\end{eqnarray}
with 
\begin{eqnarray}
	\mf{K}(\mf{q}, \mf{v})&=&-\frac{\partial \mf{F}(\mf{q}, \mf{v})}{\partial \mf{q}}\frac{\partial\dot{\mf{q}}^{\theta_{vq}}}{\partial\mf{v}},\\
	\mf{D}(\mf{q}, \mf{v})&=&-\frac{\partial \mf{F}(\mf{q}, \mf{v})}{\partial \mf{v}},
	\label{eq:dissipation_matrix}
\end{eqnarray}
the stiffness and damping matrix respectively.

We can now write the gradient of $\mf{m}(\mf{v})$ in Eq. (\ref{eq:m_taylor_expansion}) as
\begin{eqnarray}
	\nabla\mf{m}^* = \mf{A} + \mf{E}^* - dt\nabla\mf{G}
\end{eqnarray}
with 
\begin{eqnarray}
	\mf{A}=\mf{M}+ dt^2\theta_q\theta_{qv}\mf{K}+dt\theta_v\mf{D},
\end{eqnarray}
a positive definite matrix since we assumed the gradients of $\mf{F}$ to be
positive definite.

With these definitions the Taylor expansion in Eq. (\ref{eq:m_taylor_expansion})
becomes
\begin{multline}
	\nabla\mf{m}^*(\mf{v}-\mf{v}^*) =\\
	\mf{A}(\mf{v}-\mf{v}^*) + \mf{E}^*(\mf{v}-\mf{v}^*) - dt\,\nabla\mf{G}\,(\mf{v}-\mf{v}^*).
\end{multline}

In the absence of impacts $\Vert\mf{v}-\mf{v}^*\Vert=\mathcal{O}(dt)$. Therefore
$\mf{E}^*(\mf{v}-\mf{v}^*)=\mathcal{O}_E(dt^3)$ and
$dt\nabla\mf{G}(\mf{v}-\mf{v}^*)=\mathcal{O}_G(dt^2)$. Then the Taylor expansion
\begin{eqnarray}
	\mf{A}(\mf{v}-\mf{v}^*) + \mathcal{O}_E(dt^3) + \mathcal{O}_G(dt^2) = \mf{J}^T\mf{\bgamma}
\end{eqnarray}
is a first order approximation of the original momentum balance in Eq. (\ref{eq:v_update}).

\end{IEEEproof}

\subsection{Contact Resolution}
\label{sec:contact_constrained_velocities}

Once free motion velocities $\mf{v}^*$ are solved from Eq.
(\ref{eq:vstar_definition}), the second stage consists in solving for the
impulses $\bgamma$ that satisfy the linearized momentum in Eq.
(\ref{eq:momentum_linearized}) subject to contact constraints.

We now replace our expansion of $\mf(\mf{v})$ to 
obtain an approximation of Eq. (\ref{eq:v_update})
\begin{eqnarray}
	\mf{A}(\mf{v}-\mf{v}^*)= \mf{J}^T\mf{\gamma}
	\label{eq:momentum_linearized}
\end{eqnarray}
with
\begin{eqnarray}
	\mf{A}&=&\mf{M}+\theta_q\theta_{qv} dt^2\,\mf{K}+\theta_v dt\,\mf{D},
	\label{eq:expression_for_A}\\
	\mf{K}&=&-\frac{\partial \mf{F}}{\partial \mf{q}}\Bigr|_{(\mf{q}^{\theta_q}, \mf{v}^{\theta_v})},\\
	\mf{D}&=&-\frac{\partial \mf{F}}{\partial \mf{v}}\Bigr|_{(\mf{q}^{\theta_q},
	\mf{v}^{\theta_v})}.
	\label{eq:dissipation_matrix}
\end{eqnarray}
where the only assumption in Eq. (\ref{eq:expression_for_A}) is that the map
$\mf{N}(\mf{q})$ is the identity \GreenHighlight{Recommend never introducing N
if it is always the identity} which is commonly the case for degrees of freedom
used to model compliant elements. When $\mf{F}(\mf{q}, \mf{v})$ is linear the
stiffness matrix $\mf{K}$ and the damping matrix $\mf{D}$ are constant and the
two stage process described by Eqs. (\ref{eq:vstar_definition}) to
(\ref{eq:momentum_linearized}) is exact and solves the original problem stated
by Eq. (\ref{eq:v_update}).




A typical example of constant stiffness and damping is that of a spring-damper
model at a joint. In this case these matrices only contribute positive diagonal
elements to matrix $\mf{A}$.

An example of complex stiffness and damping matrices is when using a Finite
Element Model (FEM) of large soft body deformations. In this case $\mf{K}$ and
$\mf{D}$ are large sparse, positive definite, matrices.



Notice that, in the
absence of constraint impulses, the next step velocity is $\mf{v}=\mf{v}^*$ and
it is computed with the order of accuracy of the $\theta\text{-method}$. 



We also
expect to recover the order of accuracy of the $\theta\text{-method}$ when
constraints do not produce work. For instance, for a rolling wheel the stiction
constraint does not dissipate energy since the slip velocity is zero even though
the friction force is non-zero, i.e. $\bgamma_t\cdot\mf{v}_t = 0$. Moreover, we
will see in Section XXX that the symplectic midpoint rule conserves energy
exactly during rolling.\RedHighlight{Alejandro: Add proper reference to the
billiard balls example in the numerical examples section.}

Our convex approximation of contact requires $\mf{A}\succ 0$. Therefore only
compliant and dissipation elements that lead to positive definite matrices can
be included in Eq. (\ref{eq:expression_for_A}). This is the case for joint
spring-damper models and deformable models. It is not the case for Coriolis and
gyroscopic terms. These terms can still be treated implicitly in Eqs.
(\ref{eq:vstar_definition}) and (\ref{eq:qstar_definition}), but their
derivative is assumed to be zero in Eq. (\ref{eq:dissipation_matrix}), an
approximation only accurate to first order. \GreenHighlight{What does "treated
implicitly" mean? In what way is their derivative is zero in
\ref{eq:dissipation_matrix}? When you introduced $F(q,v)$ you said that it
already includes gyroscopic and Corolis terms, but this discussion seems to
about how to add them, as if they weren't already there...  }

