\documentclass[12pt]{article}

\input{packages_and_custom_commands.tex}

\usepackage[margin = 1in]{geometry}
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ...
\usepackage{amsmath}
%\setlength{\parindent}{0pt}

\begin{document}

\author{Alejandro Castro, Frank Permenter, Xuchen Han}
\title{Response to the Reviewers \\
  \large An Unconstrained Convex Formulation of Compliant Contact}
\maketitle

We thank the reviewers for their valuable time, comments and overall positive
assessment of our work. As suggested by Reviewer 1 and the Associate Editor, we have
significantly reorganized the contents. In particular, now sections II
and III of the paper are organized with the attempt to make the reading process
more linear. We also identified text repetition and removed it. On top of this
newly organized text, we addressed the comments from the reviewers. In the
following, the comments of the reviewers are reproduced here in
\textcolor{blue}{blue} for the reviewers' convenience, followed by our
responses. We numbered the comments/questions from the reviewers in this
document and marked edits in the revised manuscript with \textcolor{red}{Q\#},
where \textcolor{red}{\#} corresponds to the question number. All revised text
and responses are marked in \textcolor{red}{red}. We hope that the reviewers
find our edits acceptable and the revised manuscript suitable for publication.

\input{reviewer_1.tex}
\input{reviewer_4.tex}
\input{reviewer_8.tex}

\bibliographystyle{./IEEEtran/IEEEtran}
\bibliography{sap_paper}

\end{document}
