\section{Test Cases}
\label{sec:test_cases}

We evaluate the robustness, accuracy, and performance of our method in a number
of simulation tests. All simulations are carried out in a system with 24 2.2
GHz Intel Xeon cores (E5-2650 v4) and 128 GB of RAM, running Linux. For all
solvers below however, all tests are run in a single thread.

For all of our simulations, unless otherwise specified, our model uses
$\beta=1.0$ and $\sigma=10^{-3}$ for the regularization parameters in Eq. (\ref{eq:normal_regularization}) and Eq. (\ref{eq:slip_time_scale}), respectively.

For SAP, tolerances for the stopping criteria are set to
$\varepsilon_a=10^{-16}$ and $\varepsilon_r=10^{-5}$. While the other solvers we use below use different stopping criteria, we use the same performance and accuracy metrics for all of them in order to perform a fair comparison.

\subsection{Performance Comparisons Against Other Solvers}
\label{sec:about_solvers}

We evaluate commercial software Gurobi and Mosek to solve our convex
formulation of contact. Both solvers can solve a large variety of optimization
problems and are considered the industry standard. However, Mosek is not robust
for these problems and does not always find a solution. Therefore, we only present
comparison results with Gurobi.

As an open source option, we evaluate the Geodesic interior-point methods (IPM)
from \cite{bib:permenter2020}. Geodesic IPMs, in contrast with primal-dual
IPMs, do not apply Newton's method to the central-path conditions directly.
Instead, they use geodesic curves that satisfy the complementarity slackness
condition. This method has proven to warm start effectively when applied to
Model Predictive Control Problems \cite{bib:permenter2020} and therefore we are
interested in evaluating its performance for solving our convex formulation of
contact. Moreover, we use the Geodesic IPM solver's supernodal algebra described
in Section \ref{sec:problem_sparsity} in our implementation of SAP. Using the
same algebra for SAP and Geodesic IPM allows us to perform a fair comparison between
the two solvers; performance differences can be attributed to the properties
inherent to the optimization method instead of the differences in the linear
algebra operations.

For performance comparisons, we use the steady clock from the STL
\verb+std::chrono+ library to measure wall-clock time for SAP and Geodesic IPM.
For Gurobi we access the \verb+Runtime+ property reported by Gurobi. Notice this
is somewhat unfair to SAP and Geodesic IPM since Gurobi's reported time does not
include the cost of the initial setup.

\subsection{Spring-Cylinder}
\label{sec:spring_cylinder}
\input{spring_cylinder.tex}

\subsection{Clutter}
\label{sec:clutter}
\input{clutter.tex}

\subsection{Slip Control}
\label{sec:slip_control}
\input{slip_control.tex}

\subsection{Dual Arm Manipulation}
\label{sec:dual_arm}
\input{dual_arm.tex}
